---
title: "Cat & Dog Names - Data Wrangling v1"
author: "Stephanie Huang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SETUP - A MESS ATM - TO CLEAN UP!
*NOTE - THIS CODE IS NOT EFFICIENTLY WRITTEN BUT IS MEANT TO SHOW EACH PROCESS IN A STEP BY STEP MANNER!*

https://www.r-bloggers.com/2019/04/how-to-filter-in-r-a-detailed-introduction-to-the-dplyr-filter-function/

### Load in Packages

```{r}
library(readr) #for read_csv
library(dplyr) #data wrangling e.g. bind_rows %>%
library(stringr) #working with strings e.g. str_replace
library(tidyr) #pivot_longer

```

### Batch read in the data

This code is for reading in multiple csv files at once. This way you don't have to call each one individually. Here we only show 2 files but you can use the same code to read in 100 or 1000 etc.

1.  Set working directory as source file location. *Session \> Set Working Directory \> To source file location*

2.  Determine the csv folder location (relative to script file location) & assign to `mydir`. Then list the files using `list.files` & read in anything with the .csv extension. Need `full.names = TRUE` to get proper location?

3.  Batch read in csv files using the `SetNames` will add the file name to the dataframe, you can read it in with just `lapply(myfiles, read.csv)` but it won't have the original csv file retain as an extra column.

Note if you wanted to read in excel files you would load the *readxl* package and change the pattern `"*.xlsx"` to & the function to `read_excel`

Note to self! It seems that read.csv causes the extra space & extra column issue?

```{r}
#1 - Set working directory to source file location

#2 - Determine directory & create a list of the files
mydir = "input/animals" #data file location relative to current
myfiles = list.files(path=mydir, pattern = "*.csv", full.names = TRUE) # lists out csv files

#3 - Batch read in using lapply
df_list <- setNames(lapply(myfiles, read_csv), myfiles)

```

### Combine the individual files into one master file

```{r}
cdf <- bind_rows(df_list, .id="file_id")

glimpse(cdf)
summary(cdf)

```

## Tidy up data file

### Remove blank column,

```{r}
#1 - remove the random blank column
cdf_1 <- cdf %>% select(!(blank))

glimpse(cdf_1)


#2 - rename column names
cdf_2 <- cdf_1 %>% 
  rename("type" = "animaltype",
         "age" = "AgeYears")

glimpse(cdf_2)

```

### Replacing AND to & with `str_replace`

<https://sparkbyexamples.com/r-programming/replace-string-with-another-string-in-r/> <https://sparkbyexamples.com/r-programming/r-str-replace-to-replace-matched-patterns-in-a-string/>

```{r}
unique(cdf_2$colour)

cdf_3 <- cdf_2 %>%
  mutate(colour = str_replace_all(colour, "AND|and|And|&|,|/|\\.", " & "))

unique(cdf_3$colour)

cdf_4 <- cdf_3 %>%
  mutate(across(c(1:7), str_squish)) #removes extra spaces aka squishes down 

unique(cdf_4$colour)

```

```{r}
#Another way to look at it
cdf %>%
  count(colour) %>%
  slice(1:10)

cdf_4 %>%
  count(colour) %>%
  slice(1:10)
```

```{r}
unique(cdf_4$type)

cdf_5 <- cdf_4 %>%
  mutate(across(c(2:7), str_to_title))  #captilizes only the 1st letter of the word


unique(cdf_5$type)
unique(cdf_5$colour)

colours_b4 <- cdf_5 %>%
  count(colour) 

cdf_6 <- cdf_5 %>%
  mutate(colour = str_replace_all(colour, c(
    "Blackwhite" = "Black & White",
    "Blacktan" = "Black & Tan",
    "Tanwhite" = "Tan & White",
    "Gray" = "Grey",
    "Balck" = "Black")))

colours <- cdf_6 %>%
  count(colour) %>%
  slice(1:600)

unique(cdf_6$colour)


```


#Cat Breed Recode
```{r}

cdf_6 %>% 
  filter(type == "Cat") %>%
  count(breed, sort = TRUE) %>%
  slice_max(n, n = 20)


cdf_7 <- cdf_6 %>%
  mutate(breed = str_replace_all(breed, c(
    "Domsh" = "Domestic Short Hair",
    "Dommh" = "Domestic Medium Hair",
    "Domlh" = "Domestic Long Hair",
    "Britshcat" = "British Cat",
    "Russianbl" = "Russian Blue")))

cat_breeds <- cdf_7 %>% 
  filter(type == "Cat") %>%
  count(breed, sort = TRUE) 




```

```{r}
unique(cdf_7$type)
unique(cdf_7$sex)
unique(cdf_7$id)

cdf_8 <- cdf_7 %>%
  mutate(type = recode(type, "D" = "Dog")) %>% 
  mutate(region = recode(file_id, 
                     "input/animals/golden-plains-animal-registrations_v2.csv" =
                     "Golden Plains",
                     "input/animals/sunshinecoast_qld_animals_registered_v2.csv" =
                     "Sunshine Coast"), .before = suburb)

unique(cdf_8$type)
glimpse(cdf_8)  

```

```{r}

cdf_9 <- cdf_8 %>% 
  mutate(animal_id = row_number(), .after = file_id)

glimpse(cdf_9)

```

### Export out the file

```{r}
#write_csv(cdf_9, "cleaned_catdog_data_v2_cat.csv")

```

## Pivot to long form

Not actually that relevant for this dataset but is useful in other cases. Here is some demo code on how to do it with out data.

```{r}
long_df <- cdf_9 %>%
  select(!file_id) %>%
  pivot_longer(cols = c('name':'suburb'),
               names_to = 'measure',
               values_to = 'value')

str(long_df)
```

```{r}
#write_csv(long_df, "long_cleaned_catdog_data_v1.csv")
```
